Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels

### 2016-07-13 PALSE/talk_Kyoto.tex
scaling theory in the middle of the slide
forgotten strain.pgf extention of the electrostatic blobs => tau => E_c => A

\begin{frame}{Polyelectrolytes physics}
	\structure{Can we measure counterion condensation from rheology?}

	\begin{itemize}
		\item 1 dissociated counterion every \alert{$A$} monomers
		\begin{itemize}\item large $A$, more condensed\end{itemize}
		\item poor solvent: reduced temperature $\alert{\tau} = 1 -T/\Theta$
		\begin{itemize}\item \ce{PIm+F-} non soluble at \SI{100}{\celsius} $\Rightarrow\tau>0.2$ \end{itemize}
		\smallskip
		\item extension parameter $\alert{B} = \frac{\text{extended length}}{\text{diameter}}$ of electrostatic blob
		\smallskip
		\item from scaling theory: $B \sim A^{4/3} \tau$\hfill \textit{\scriptsize Rubinstein Macromol. 1996}
	\end{itemize}

	\tikzsetnextfilename{scales}%
	\input{fig_presentation/scales.pgf}
	\begin{center}
	\structure{Can we measure $B$ from rheology?}
	\end{center}
\end{frame}

>>>>

\begin{frame}{Polyelectrolytes physics}
	\vspace{\baselineskip}
	\structure{Can we measure counterion condensation from rheology?}

	\begin{itemize}
		\item 1 dissociated counterion every \alert{$A$} monomers
		\begin{itemize}\item large $A$, more condensed\end{itemize}
		\item poor solvent: reduced temperature $\alert{\tau} = 1 -T/\Theta$
		\begin{itemize}\item \ce{PIm+F-} non soluble at \SI{100}{\celsius} $\Rightarrow\tau>0.2$ \end{itemize}
	\end{itemize}
	\alert{Scaling theory}\hfill \textit{\scriptsize Rubinstein Macromol. 1996}
	\tikzsetnextfilename{scales}%
	\input{fig_presentation/scales.pgf}

	\begin{itemize}
		\item extension parameter $\alert{B} = \frac{\text{extended length}}{\text{diameter}}$ of electrostatic blob
		\smallskip
		\item scales $\geq D$ depend only on $B \sim A^{4/3} \tau$
	\end{itemize}

	\begin{center}
	\structure{Can we measure $B$ from rheology?}
	\end{center}
\end{frame}

\begin{frame}<1>[label=endlinear]{Stretching a polyelectrolyte chain}
	\vspace{\baselineskip}
	\begin{itemize}
		\item Unfold largest scales first\hfill \textit{\scriptsize Pincus Macromol. 1976}
	\end{itemize}

	\tikzsetnextfilename{scales_stretch}%
	\begin{tikzpicture}
	\node[inner sep=0](scales){\input{fig_presentation/scales.pgf}};
	\draw[->, ultra thick, ilmcolor] (scales.south east) ++(0,-0.3em) -- +(-0.33\textwidth,0) node[midway, below] {linear};
	\draw[->, ultra thick,ilmorange] (scales.south) ++(0,-0.3em) -- +(-0.35\textwidth,0) node[midway, below] {non linear};
	\end{tikzpicture}

	\pause
	\begin{itemize}
		\item narrow linear regime $\Rightarrow R_0\approx r_\mathrm{scr} \Rightarrow \left(\frac{B}{B_0}\right)^3 = 1 + \frac{B}{B_s}$
		\begin{itemize}
			\item \alert{$B_s$} screening from \ce{-PO_3^{2-}} heads\tikzsetnextfilename{head}
				\begin{tikzpicture}
					\let\mrad\relax%
					\newlength\mrad%
					\setlength{\mrad}{1ex}%
					%face
					\input{fig_presentation/head.pgf}
					%% horns
					\draw[every node/.style={shape=circle,inner sep=0.2\mrad,shading=radial,inner color=white!50!ilmcolor,outer color= ilmcolor!90!black}] (emoticon.70) -- ++(70:0.4\mrad) node{} (emoticon.110) -- ++(110:0.4\mrad) node{};
				\end{tikzpicture}
			\item if $B \ll B_s$ screening from uncondensed counterions \[B = B_0 \sim bc \left(\frac{k_\mathrm{B}T}{G}\right)^{2/3}\]
			\item if $B \gg B_s$ screening from heads: $B = \left(B_0^3/B_s\right)^{1/2} > B_0$
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}<1>[label=strainsweep]{Strain sweep}
	\vspace{0.5\baselineskip}
	\tikzsetnextfilename{strain}%
	\input{fig_presentation/strain.pgf}

	\vspace{-0.5\baselineskip}
	\begin{overprint}%
	\onslide<-2>%
	\begin{itemize}
		\item narrow linear regime $\gamma_0\ll 1$
		\item<2> extension parameter increases with condensation
	\end{itemize}
	\onslide<3->%
	\begin{itemize}
		\item bonding energy increases with $B$
		\item<4> measure of counterion condensation $A$
%		\begin{description}
%			\item[\ce{PPyr+I-}] $10$ uncondensed counterions per polymer
%			\item[\ce{PIm+Br-}] a single dissociated counterion every $9$ polymers
%		\end{description}
	\end{itemize}
	\end{overprint}
\end{frame}

\againframe<2>{endlinear}
\againframe<2>{strainsweep}

\begin{frame}{Extension of the electrostatic blobs}
\vspace{\baselineskip}

\tikzsetnextfilename{stretch}%
	\tikzsetnextfilename{stretch}%
	\input{fig_presentation/stretch.pgf}
	\[E_\mathrm{stretch} = \text{area increment}(\gamma,B) \times \text{surface tension}(\tau,B)\]

	\begin{itemize}
		\item fluidisation = rupture $\Rightarrow E_\mathrm{stretch}(\gamma_c) = E_\text{head-body}$
		\begin{itemize}
			\item \structure{$\tau(E_\text{head-body}, B, \gamma_c)$}
		\end{itemize}
		\item ionic bond energy depends on local $\epsilon_r$%\\\hfill\textit{\scriptsize Rehm Chem Soc Rev. 2010}
		\begin{itemize}
			\item in water $E_\text{head-body}\approx 2k_\mathrm{B}T$
			\item large $B\Rightarrow$ low $\epsilon_r$ $\Rightarrow$ high $E_\text{head-body}$
		\end{itemize}
		\item smallest $B$ is \ce{PPyr+I-} $\Rightarrow \textcolor{ilmcolor}{\tau = 0.4\pm0.07}, \Theta\approx\SI{220}{\celsius}$
	\end{itemize}
\end{frame}

\againframe<3-4>{strainsweep}
